---
title: "ST558_HW4 - Lee Worthington"
format: html
editor: visual
---

## Setup environment

```{r}
#| warning: FALSE
#| message: FALSE

library(httr)
library(jsonlite)
library(magrittr)
library(purrr)
library(dplyr)
```

## Task 1: conceptual questions

### Question 1 - What is the purpose of the lapply() function? What is the equivalent purrr function?

> lapply() is the baseR function that applies a function to every element in a list individually. For instance, you could use lapply() to find the mean of the values in element 1, then the mean of the values in element 2, etc. The equivalent purrr function is map()

### Question 2 - code to use lapply() and cor()

```{r}
#| eval: false

lapply(
  X = my_list,
  FUN = cor,
  method = 'kendall'
)

```

### Question 3 - what are the two advantages of using purrr functions isntead of the baseR apply family?

> purrr functions:
>
> -   Have helper functions that improve functionality/readability, such as the map\_\*() functions
> -   There is much greater consistency between purrr functions compared to the apply family in terms of arguements, output, options, etc

### Question 4 - what is a side effect function?

> Side effect functions do not try to change/transform the data being passed to them, rather they try to produce something (side effects) from the data. Examples owuld be print()/plot() which generate output from the data without altering it

### Question 5 - why can you name a variable sd in a function and not cause any issues with the sd function?

> The variable sd defined in the function only exists in the temporary environment and scope of the function, so it only exists while the function is running and therefore cannot cause any errors with the global environment.

## Task 2: writing R functions

### Question 1 - writing a function to calculate RMSE

```{r}
#| eval: true

# define function
getRMSE <- function(observation_vector, prediction_vector, ...){
  
  # calculate differences in vector
  differences <- observation_vector - prediction_vector
  
  # square the difference
  squared_differences <- differences^2
  
  # calculate the mean of the squared diffs, with optional argument
  avg_differences <- mean(squared_differences, ...)
  
  # take the root
  rmse <- sqrt(avg_differences)
  
  # return rmse
  return(rmse)
}
```

### Question 2 - generate fake data and test RMSE function

```{r}
#| eval: true

# set seed for reproducibility
set.seed(10)

# generate fake data
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

# test getRMSE function on original data
rmse_original <- getRMSE(
  observation_vector = resp,
  prediction_vector = pred
)

# add NAs to response data
resp_na <- resp
resp_na[c(3, 88)] <- NA_real_

# test getRMSE function on NA induced data
rmse_NAs_no_arguement <- getRMSE(
  observation_vector = resp_na,
  prediction_vector = pred
)

rmse_NAs_w_arguement <- getRMSE(
  observation_vector = resp_na,
  prediction_vector = pred,
  na.rm = TRUE
)

# print results
rmse_original
rmse_NAs_no_arguement
rmse_NAs_w_arguement
```

### Question 3 - writing a function to calculate MAE

```{r}
#| eval: true

# define function
getMAE <- function(observation_vector, prediction_vector, ...){
  
  # calculate differences in vector
  differences <- observation_vector - prediction_vector
  
  # get the absolute differences
  absolute_differences <- abs(differences)
  
  # calculate the mean of the absolute diffs, with optional argument
  mae <- mean(absolute_differences, ...)
  
  # return mae
  return(mae)
}
```

### Question 4 - generate fake data and test MAE function

```{r}
#| eval: true

# set seed for reproducibility
set.seed(10)

# generate fake data
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

# test getMAE function on original data
mae_original <- getMAE(
  observation_vector = resp,
  prediction_vector = pred
)

# add NAs to response data
resp_na <- resp
resp_na[c(3, 88)] <- NA_real_

# test getMAE function on NA induced data
mae_NAs_no_arguement <- getMAE(
  observation_vector = resp_na,
  prediction_vector = pred
)

mae_NAs_w_arguement <- getMAE(
  observation_vector = resp_na,
  prediction_vector = pred,
  na.rm = TRUE
)

# print results
mae_original
mae_NAs_no_arguement
mae_NAs_w_arguement
```

### Question 5 - create a wrapper for MAE and RMSE functions

```{r}
#| eval: true

# define metrics function with default statistics
getMetrics <- function(observation_vector, prediction_vector, statistics = c('RMSE', 'MAE'), ...){
  
  # check to make sure inputs are correct type
  if(!is.vector(observation_vector) | !is.atomic(observation_vector) | !is.numeric(observation_vector)){
    stop('observation_vector is not an atomic numeric vector')
  }
  if(!is.vector(prediction_vector) | !is.atomic(prediction_vector) | !is.numeric(prediction_vector)){
    stop('prediction_vector is not an atomic numeric vector')
  }
  
  # create a list to store the statistics
  output_statistics <- list()
  
  # calculate statistics conditionally
  if('RMSE' %in% statistics){
    output_statistics$RMSE <- getRMSE(observation_vector, prediction_vector, ...)
  } 
  if ('MAE' %in% statistics){
    output_statistics$MAE <- getMAE(observation_vector, prediction_vector, ...)
  }
  
  # calculate everything if no valid inputs given
  if (length(output_statistics) == 0) {
    output_statistics$RMSE <- getRMSE(observation_vector, prediction_vector, ...)
    output_statistics$MAE <- getMAE(observation_vector, prediction_vector, ...)
  }
  
  # return metrics
  return(output_statistics)
}

```

### Question 6 - generate fake data and test wrapper

```{r}

# set seed for reproducibility
set.seed(10)

# generate fake data
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

# call wrapper and generate stats
mae <- getMetrics(
  observation_vector = resp,
  prediction_vector = pred,
  statistics = 'MAE'
)
rmse <- getMetrics(
  observation_vector = resp,
  prediction_vector = pred,
  statistics = 'RMSE'
)
all_statistics <- getMetrics(
  observation_vector = resp,
  prediction_vector = pred,
  statistics = c('MAE', 'RMSE')
)

# add NAs to response data
resp_na <- resp
resp_na[c(3, 88)] <- NA_real_

# call wrapper and generate stats on missing data
mae_NAs <- getMetrics(
  observation_vector = resp_na,
  prediction_vector = pred,
  statistics = 'MAE',
  na.rm = TRUE
)
rmse_NAs <- getMetrics(
  observation_vector = resp_na,
  prediction_vector = pred,
  statistics = 'RMSE',
  na.rm = TRUE
)
all_statistics_NAs <- getMetrics(
  observation_vector = resp_na,
  prediction_vector = pred,
  statistics = c('MAE', 'RMSE'),
  na.rm = TRUE
)

# print no NA data
mae
rmse
all_statistics

# print NA data
mae_NAs
rmse_NAs
all_statistics_NAs

# pass incorrect data structure (should return an error)
try({
  mae <- getMetrics(
    observation_vector = as.data.frame(c('1', 'a')),
    prediction_vector = as.data.frame(c('1', 'a')),
    statistics = 'MAE'
  )
})
```

## Task 3: querying APIs

### Question 1 - get API data

```{r}
# set api key
api_key <- 'f07771c038d24e65aeaf85752af8bb4a'

# set url for API
url <- 'https://newsapi.org/v2/everything?q=tesla&from=2024-05-29&sortBy=publishedAt&apiKey=f07771c038d24e65aeaf85752af8bb4a'

# get the data form the API
url_data <- httr::GET(url)
```

### Question 2 - pull the article data

```{r}
# call url_data
articles <- url_data |>
  
  # parse the url_data as text
  httr::content(as = 'text') |>
  
  # convert the JSON text into a dataframe format
  fromJSON(flatten = TRUE, simplifyDataFrame = TRUE) |>
  
  # pull specified element named articles, which is a list of 10 elements
  pluck('articles') |>
  
  # grab the content column
  select(content)

# print results
head(articles$content, 5)
```

### Question 3 - api function

```{r}

# create function
query_api <- function(subject, start_date, api_key){
  
  # set base url
  url <- 'https://newsapi.org/v2/everything'
  
  # make a request to the API
  url_data <- httr::GET(url, query = list(
    q = subject,
    from = start_date,
    sortBy = 'publishedAt',
    apiKey = api_key
  ))
  
  # parse the article data
  content <- url_data |>
    
    # parse the url_data as text
    httr::content(as = 'text') |>
    
    # convert the JSON text into a dataframe format
    fromJSON(flatten = TRUE, simplifyDataFrame = TRUE) |>
    
    # pull specified element named articles, which is a list of 10 elements
    pluck('articles') |>
    
    # grab the content column
    select(content)
  
  # return results
  return(content)
}

# test function (dont forget time limit)
head(query_api('science', '06-06-2024', 'f07771c038d24e65aeaf85752af8bb4a')$content, 5)
head(query_api('politics USA', '06-06-2024', 'f07771c038d24e65aeaf85752af8bb4a')$content, 5)
```